---
title: "MATH 660 SP24 Homework 2 on ggplot2"
author: "Dontese Cocchio"
date: "Due 2023-03-04 11:59PM (Monday)"
output:
  html_document: default
  toc: true
theme: united
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this homework we deal with datasets from the following link:

<https://github.com/plotly/datasets>

Please use the functions in the `tidyverse`, i.e., the functions introduced in the textbook after Chapter 4,
to get the full score.

## Layers

## Question 1

We use the wind rose data.

<https://github.com/plotly/datasets/blob/master/wind_rose.csv>

Read the input and "lengthens" the data so that it is reshaped into long format.
The reshaped table will be called `wind_rose_long_data`.

Please filter out the directions "N", "NE", "E", "SE", "S", "SW", "W", and "NW",
and create the figure shown below.

Hint: For barplot, you may want to use `geom_bar()`. You may want to use the
`ggarrange()` function in the `ggpubr` package. You may use `theme_minimal()`
for the theme.

```{r Q1}
library(tidyverse)
library(ggpubr)
wind_rose_wide_data <- read_csv("https://raw.githubusercontent.com/plotly/datasets/master/wind_rose.csv")
############################################################
# sorting the data
NW <- filter(wind_rose_long_data, r == "NW")
N <- filter(wind_rose_long_data, r == "N")
NE <- filter(wind_rose_long_data, r == "NE")
E <- filter(wind_rose_long_data, r == "E")
SE <- filter(wind_rose_long_data, r == "SE")
S <- filter(wind_rose_long_data, r == "S")
SW <- filter(wind_rose_long_data, r == "SW")
W <- filter(wind_rose_long_data, r == "W")

# individual plots
plot1 <- ggplot(NW, aes(speed, frequency)) +
  geom_bar(stat="identity", fill="red") +
  ggtitle("NW") +
  scale_y_continuous(breaks = seq(0, 2.5, by = 1), limits = c(0, 2.5)) +
  theme_minimal()
plot2 <- ggplot(N, aes(speed, frequency)) +
  geom_bar(stat="identity", fill="red") +
  ggtitle("N") +
  scale_y_continuous(breaks = seq(0, 2.5, by = 1), limits = c(0, 2.5)) +
  theme_minimal()
plot3 <- ggplot(NE, aes(speed, frequency)) +
  geom_bar(stat="identity", fill="red") +
  ggtitle("NE") +
  scale_y_continuous(breaks = seq(0, 2.5, by = 1), limits = c(0, 2.5)) +
  theme_minimal()
plot4 <- ggplot(E, aes(speed, frequency)) +
  geom_bar(stat="identity", fill="red") +
  ggtitle("E") +
  scale_y_continuous(breaks = seq(0, 2.5, by = 1), limits = c(0, 2.5)) +
  theme_minimal()
plot5 <- ggplot(SE, aes(speed, frequency, )) +
  geom_bar(stat="identity", fill="red") +
  ggtitle("SE") +
  scale_y_continuous(breaks = seq(0, 2.5, by = 1), limits = c(0, 2.5)) +
  theme_minimal()
plot6 <- ggplot(S, aes(speed, frequency)) +
  geom_bar(stat="identity", fill="red") +
  ggtitle("S") +
  scale_y_continuous(breaks = seq(0, 2.5, by = 1), limits = c(0, 2.5)) +
  theme_minimal()
plot7 <- ggplot(SW, aes(speed, frequency)) +
  geom_bar(stat="identity", fill="red") +
  ggtitle("SW") +
  scale_y_continuous(breaks = seq(0, 2.5, by = 1), limits = c(0, 2.5)) +
  theme_minimal()
plot8 <- ggplot(W, aes(speed, frequency)) +
  geom_bar(stat="identity", fill="red") +
  ggtitle("W") +
  scale_y_continuous(breaks = seq(0, 2.5, by = 1), limits = c(0, 2.5)) +
  theme_minimal()
plot9 <- ggplot(data=NULL) + 
  theme_void()  
# combining plots
ggarrange(plot1, plot2, plot3, plot8, plot9, plot4, plot7, plot6, plot5, nrow=3, ncol=3)
############################################################
```

```{r A1}
wind_rose_long_data <- pivot_longer(wind_rose_wide_data, `0-1`:`6+`, names_to = "speed", values_to = "frequency")
wind_rose_long_data
```


## Question 2

We use the tesla stock price data.

Before starting this task, please read the chapter on dealing with dates and times using the `lubridate` package:

> Chapter 25 Parsing dates and times:
>
> <http://rafalab.dfci.harvard.edu/dsbook/parsing-dates-and-times.html>


For each year, please get the highest `close` price and the lowest `close` price.

The table `max_and_min_close_per_year` will have 4 obs. of 3 variables. The column names are `year` for each year, `max_close` for the highest `close` price, and `min_close` for the lowest `close` price.

Please create a line chart shown below.

Hint: The color we use is `scale_color_brewer(palette = "Pastel2")` and
the theme is `theme_classic()`.

```{r Q2}
library(tidyverse)
library(RColorBrewer)
tesla_stock_price <- read_csv("https://raw.githubusercontent.com/plotly/datasets/master/tesla-stock-price.csv")
# You may notice the date format in the first row is different. Let us just set it to "2018/10/16":
tesla_stock_price[1, "date"] <- "2018/10/16"
library(lubridate)
############################################################
max_and_min_close_per_year_longer |> 
  group_by(max_or_min) |>
  ggplot(aes(year, stock_price, color = max_or_min)) +
  geom_point() +
  geom_line(aes(linetype = max_or_min)) +
  scale_color_brewer(palette = "Pastel2")
  theme_classic() +
  theme(  panel.background = element_rect(fill='transparent'),
         plot.background = element_rect(fill='transparent', color=NA),
         panel.grid.major = element_blank(),
         panel.grid.minor = element_blank())
  # still cant remove the grid in the background.
############################################################
```

```{r A2}
max_and_min_close_per_year = tesla_stock_price |>
  mutate(year = year(ymd(date))) |>
  group_by(year) |>
  summarize(max_close = max(close),
            min_close = min(close))
max_and_min_close_per_year
max_and_min_close_per_year_longer <- max_and_min_close_per_year |>
  pivot_longer(max_close:min_close, names_to = "max_or_min", values_to = "stock_price")
```


## Question 3

We use the Tesla stock price data.

What is the Average Daily Trading Volume (ADTV) of the Tesla stock per month in year 2017?

What I expect: The tibble `ADTV_in_2017` will have 12 obs. of 2 variables. The column names are `month` for each month and `ADTV` for the mean volume per month.

Hint: The y axis is log10 scaled. Every month on the x axis needs to be labeled.
For accessibility, the color palette `viridis` is color blind-friendly and is from
the `viridis` package. The plot is a `geom_dotplot`. The theme is `theme_classic`.
You may need to specify that the color depends on `month` as a factor and 
use a discrete palette by specifying `discrete = TRUE` when using the `viridis` color scheme.

```{r Q3}
library(tidyverse)
tesla_stock_price <- read_csv("https://raw.githubusercontent.com/plotly/datasets/master/tesla-stock-price.csv")
# You may notice the date format in the first row is different. Let us just set it to "2018/10/16":
tesla_stock_price[1, "date"] <-  "2018/10/16"
library(lubridate)
############################################################
library(viridis)
daily_volume_in_2017 |>
  ggplot(aes(factor(month), volume, fill = factor(month))) +
  geom_dotplot(binaxis = "y", stackdir = "center", position = 'identity', 
               binwidth = NULL, dotsize = .7) +
  scale_x_discrete(breaks = 1:12, labels = 1:12) +
  scale_y_continuous(trans = "log10") +
  labs(x = "Month", y = "Volume") +
  scale_fill_viridis(discrete=TRUE) +
  theme_classic()
  # can't quite get the dots to stack as they do in your plot
############################################################
```


```{r A3}
daily_volume_in_2017 = tesla_stock_price |>
  mutate(date = ymd(date), year = year(date), month = month(date)) |>
  filter(year == 2017)
daily_volume_in_2017
```


## Question 4

In this question we analyze a dataset of United States cities.

<https://github.com/plotly/datasets/blob/master/us-cities-top-1k.csv>

Please find the top 10 cities by `Population`, sort by decreasing population, and only select the columns of 
`City`, `State`, and `PopulationInMillions` (= Population / 1,000,000).


Hint: 
  For the color, please use `scale_color_viridis` and add `option="magma"` for the color palette and `trans = "log"` for color in log scale.
  To unify the legends with both color and size information, please specify the same limits and breaks `limits=c(0.5, 16), breaks = c(2,4,8))` in both `scale_color_viridis` and `scale_size_continuous`. Then you need to add `guides(color=guide_legend(), size = guide_legend())` to finally merge the legends.

```{r Q4}
library(tidyverse)
us_cities_top_1k <- read_csv("https://raw.githubusercontent.com/plotly/datasets/master/us-cities-top-1k.csv")
############################################################
us_top_10_populous_cities_with_location |>
  ggplot(aes(lon, lat, size = PopulationInMillions)) +
  geom_point() +
  geom_text(aes(label = City)) +
  scale_size_continuous(name = 'PopulationInMillions', limits = c(0.5, 16), breaks = c(2, 4, 8)) +
  scale_color_viridis(limits = c(0.5, 16), breaks = c(2, 4, 8), option = "magma", trans = "log") +
  labs(x = "Longitude", y = "Latitude") +
  guides(color = guide_legend(), size = guide_legend())

############################################################
```

```{r A4}
us_top_10_populous_cities_with_location <-  us_cities_top_1k |>
  arrange(desc(Population)) |>
  slice(1:10) |>
  mutate(PopulationInMillions = Population / 1e6) |>
  select(City, State, PopulationInMillions, lat, lon)
us_top_10_populous_cities_with_location
```


## Question 5

In this question we analyze a dataset of beers.

<https://github.com/plotly/datasets/blob/master/beers.csv>

Please print the names of beers (the column `beer`) from breweries in the top 10 most populous cities obtained in Question 4, arranged by descending order of `ounces`, and only retain the beers with `ounces` of greater than or equal to a pint (= 16 ounces).

Please create a stacked bar plot using a color blind-friendly palette (for example, "turbo" or other palettes in the viridis package).

Hint: The color palette is `scale_fill_viridis(option = "turbo")`.
For setting the x axis, use: `theme(axis.text.x = element_text(angle = 45, vjust = 0.5))`.

```{r Q5}
library(tidyverse)
beers <- read_csv("https://raw.githubusercontent.com/plotly/datasets/master/beers.csv")
# Make sure you know the column names of the table `beers`:
names(beers)
############################################################
beer_ounces %>% 
  ggplot(aes(city, n, fill = ounces)) +
  geom_bar(stat = 'identity') +
  labs(x = 'city', y = "number of beer products", fill = 'Ounces') +
  scale_fill_viridis(option = "turbo") +
  theme_minimal()
############################################################
```

```{r A5}
beer_ounces <- beers |> 
  inner_join(us_top_10_populous_cities_with_location, by = c("city" = "City")) |>
  group_by(city) |>
  count(ounces)
beer_ounces
```

